### 2019-10-29 cover_letter.tex
arXiV

\documentclass{letter}
\usepackage{txfonts}

\setlength{\textheight}{10.2in} 
\setlength{\topmargin}{-0.9in}

\signature{Dr Simon Cotter (corresponding author), Prof. Ioannis
  Kevrekidis, and Dr Paul Russell}
\address{School of Mathematics \\ Alan Turing Building \\ Oxford
  Road\\Manchester\\ M13 9PL \\ UK \\\textit{email:} simon.cotter@manchester.ac.uk}

\begin{document}

\date{}
\begin{letter}{ }

\opening{Dear Prof. Stuart and Prof. Challenor,}

Please find attached our new manuscript entitled ``Transport map accelerated adaptive importance sampling, and application to inverse problems arising from
  multiscale stochastic reaction networks''. In this paper, we expand
  on recent work on adaptive importance samplers, by incorporating the
  use of transport maps to simplify challenging posterior
  distributions, giving rise to efficient and stable importance
  sampling. The efficacy of the approach is demonstrated through
  application to inverse problems for multiscale biochemical networks,
  for which the posterior is highly concentrated close to curved lower
  dimensional manifolds.

We feel that SIAM Journal of Uncertainty
  Quantification would be the best place for this work, since the main
  two references, one by Youssef Marzouk's group and the other by my
  own group, are published and in review respectively, with this journal.
We very much look forward to hearing from you in due course regarding this paper.

\closing{Yours sincerely,}
\end{letter}

\end{document}